---
title       : Stock Forecast
subtitle    : A Shiny Web Application
author      : Mike Chu
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---
    
## Introduction
    
This application is using Rob Hyndman's BATS forecasting model
and trying to forecast the stock price 3 days in future.

Stock price is a typical **Time Series** data. BATS model is exploiting
the complex seasonal patterns in the data with exponential smoothing and
performing the forecast. In case you are interested, you could reference
this journal.

[Forecasting time series with complex seasonal patterns using exponential smoothing](http://robjhyndman.com/papers/ComplexSeasonality.pdf)

### Anything more interesting than looking into the future?

Visit the application online here now at:

### [Stock Forecast](http://mike-chu.shinyapps.io/stockForecast)

---&twocol w1:60% w2:40%

## Data Source

*** =left
This application is using R's **quantmod** package. Users will specify 
the stock symbols and the data range as source for the forecast. We will
download the closing price information from Yahoo finance webpage and
provide a time series chart for data visualization.

As with any data prediction, we should have a good data source that could
show us some patterns for the prediction. User can observe in the chart if
there are any seasonal patterns for the forecast.

*** =right
```{r quantmod, results = 'asis', comment = NA, message = F, echo = F}
library(quantmod)
input = list(symb="SPY",dates=c("2014-01-01","2014-08-13"))
data.xts <- getSymbols(input$symb, src = "yahoo", 
                       from = input$dates[1],
                       to = input$dates[2],
                       auto.assign = FALSE)
SPY <- data.xts
chartSeries(SPY, theme = chartTheme("white"), 
            type = "line", TA = NULL)
```

---
    
## Interactive Plot with NVD3.js
    
NVD3 has a very interactive line chart with view finder. You can find selective legend,
mouse-over focus to show data points at top chart and the view finder for time range
at the bottom. It is making the application much more appealing.

```{r nvd3plot, results = 'asis', comment = NA, message = F, echo = F}
library(shiny)
library(rCharts)
library(forecast)
library(reshape2)

data.df <- data.frame(index(data.xts), coredata(data.xts), stringsAsFactors=FALSE)
data.df <- data.df[,c(1, 5)]
colnames(data.df) <- c("Date", "ClosingPrice")
fit <- bats(data.df$ClosingPrice, use.parallel=FALSE)
fcast <- forecast(fit, h=3, level=95)
data.df$ForecastLowB <- rep(NA, nrow(data.df))
data.df$ForecastHighB <- rep(NA, nrow(data.df))
pred <- data.frame(Date=max(data.df$Date)+1:3,
                   ClosingPrice=rep(NA,3),
                   ForecastLowB=round(fcast$lower,2),
                   ForecastHighB=round(fcast$upper,2))
data.df <- rbind(data.df, pred)
data.melt <- melt(data.df, id.vars=1)
df <- data.melt
n1 <- nPlot(value ~ Date, group = 'variable',
            data = df,
            type = 'lineWithFocusChart'
)
n1$chart(
    color = c('blue', 'green', 'red'),
    margin = list(left = 100)
)
n1$set(
    dom = 'chart1'
)
n1$xAxis(
    axisLabel = "Date",
    tickFormat = "#!function(d) {return d3.time.format('%d-%b-%Y')(new Date( d * 86400000 ));}!#"
)
n1$x2Axis(
    tickFormat = "#!function(d) {return d3.time.format('%b-%Y')(new Date( d * 86400000 ));}!#"
)
n1$yAxis(
    axisLabel = "Stock Price",
    tickFormat = "#! function(d) {return '$' + d3.format(',.2f')(d)} !#"
)
n1$show('inline')
```

---
    
## BATS model (forecast)
    
BATS is an exponential smoothing state space model with Box-Cox transformation, ARMA errors,
Trend and Seasonal components. We can use this to forecast 3 days in horizon and with 95%
confidence level for prediction levels.

```{r bats, eval=FALSE}
fit <- bats(data.df$ClosingPrice, use.parallel=FALSE)
fcast <- forecast(fit, h=3, level=95)
```
